---
title: "8_rna_velocity"
---
```{r}
suppressPackageStartupMessages({
  library(Seurat)
  library(Signac)
  library(SeuratWrappers)
  library(BSgenome.Nfurzeri.NCBI.Nfu20140520.custom)
  library(tidyverse)
})



#import helper functions
source("C:/Users/Christopher He/OneDrive/UCSF/Singh Lab/diapause-sc-multiome/code/helper_functions.R")

#set output and figure directories
outs_dir <- "../results_all/"
fig_dir <- "../results_all/"
```
#Load Seurat Object
```{r}
so <- readRDS("../results_all/so_neurons.rds")
```


```{r}
so$barcode <- colnames(so)
so$umap.rna_1 <- so@reductions$umap.rna@cell.embeddings[,1]
so$umap.rna_2 <- so@reductions$umap.rna@cell.embeddings[,2]
so$umap.atac_1 <- so@reductions$umap.atac@cell.embeddings[,1]
so$umap.atac_2 <- so@reductions$umap.atac@cell.embeddings[,2]
so$umap.wnn_1 <- so@reductions$wnn.umap@cell.embeddings[,1]
so$umap.wnn_2 <- so@reductions$wnn.umap@cell.embeddings[,2]

# write metadata
write.csv(so@meta.data, file=paste0(outs_dir, 'metadata.csv'), quote=F, row.names=F)

# write dimesnionality reduction matrix, in this example case pca matrix
write.csv(so@reductions$pca@cell.embeddings, file=paste0(outs_dir, 'pca.csv'), quote=F, row.names=F)

# extract neighborhood graph
nn_idx <- so@neighbors$weighted.nn@nn.idx
nn_dist <- so@neighbors$weighted.nn@nn.dist
nn_cells <- so@neighbors$weighted.nn@cell.names

# save neighborhood graph
write.table(nn_idx, paste0(outs_dir, "nn_idx.txt"), sep = ',', row.names = F, col.names = F, quote = F)
write.table(nn_dist, paste0(outs_dir, "nn_dist.txt"), sep = ',', row.names = F, col.names = F, quote = F)
write.table(nn_cells, paste0(outs_dir, "nn_cells.txt"), sep = ',', row.names = F, col.names = F, quote = F)

rm(so)
gc()

# # write expression counts matrix
# library(Matrix)
# counts_matrix <- GetAssayData(seurat_obj, assay='RNA', slot='counts')
# writeMM(counts_matrix, file=paste0(out_data_dir, 'counts.mtx'))
#
# # write gene names
# write.table(
#   data.frame('gene'=rownames(counts_matrix)),file='gene_names.csv',
#   quote=F,row.names=F,col.names=F
# )
```


Note: Using python version through reticulate. scVelo and multiVelo were installed in the multiome-velocyto conda environment
```{r}
reticulate::use_condaenv("multiome-velocyto")
library(reticulate)
```

```{python}
import os
import scipy
import numpy as np
import pandas as pd
import scanpy as sc
import scvelo as scv
import multivelo as mv
import matplotlib.pyplot as plt
from pathlib import Path
```

```{python}
sc.settings.cachedir = Path('../results_all/cache')
scv.settings.verbosity = 3
scv.settings.presenter_view = True
scv.set_figure_params('scvelo')
pd.set_option('display.max_columns', 100)
pd.set_option('display.max_rows', 200)
np.set_printoptions(suppress=True)
```

## Reading in unspliced and spliced counts
```{python}
adata_rna = sc.read("../aggr_batch1-3/velocyto/velocyto_combined.loom", cache=True)

#rename barcodes to seurat format ("[barcode]-n" where n correspondes to sample)
suffixes = sum([[f"-{i}"] * n for i, n in enumerate([10177, 9774, 10510, 6575, 10736, 5993, 17082, 12245, 7928, 12501, 20000, 12563], start=1)], [])
adata_rna.obs_names = [x.split(':')[1][:-1] for x in adata_rna.obs_names]

#append "-n" suffix to each barcode
adata_rna.obs_names = [a + b for a, b in zip(adata_rna.obs_names, suffixes)]
```

```{python}
#read in metadata and pca table from seurat object
cell_meta = pd.read_csv("../results_all/metadata.csv")

#filter barcodes in seurat object
barcodes = cell_meta.barcode
adata_rna = adata_rna[adata_rna.obs_names.isin(barcodes)].copy()

#reorder metadata in order of adata
cell_meta = cell_meta.iloc[pd.Index(cell_meta['barcode']).get_indexer(adata_rna.obs_names)] 

#read in pca
pca = pd.read_csv("../results_all/pca.csv")
pca = pca.loc[cell_meta.index] # reorder pca in order of adata

#add metadata to adata
cell_meta = cell_meta.set_index('barcode')
adata_rna.obs = cell_meta


# set pca and umap
adata_rna.obsm['X_pca'] = pca.to_numpy()
adata_rna.obsm['X_umap'] = np.vstack((adata_rna.obs['umap.rna_1'].to_numpy(), adata_rna.obs['umap.rna_2'].to_numpy())).T
```


```{python}
#plot umap
sc.pl.umap(adata_rna, color=['timepoint_group'], frameon=False)
```

```{python}
# Top 2000 variable genes are used for downstream analyses.
scv.pp.filter_and_normalize(adata_rna, min_shared_counts=10, n_top_genes=2000)
```


##Preprocessing the ATAC counts (Done on wynton)
```{python}
# adata_atac = sc.read_10x_mtx('../aggr_batch1-3/filtered_feature_bc_matrix/', var_names='gene_symbols', cache=True, gex_only=False)
# adata_atac = adata_atac[:,adata_atac.var['feature_types'] == "Peaks"]
```

```{python}
# adata_atac = mv.aggregate_peaks_10x(adata_atac,
#                                     '../aggr_batch1-3/atac_peak_annotation.tsv',
#                                     '../aggr_batch1-3/feature_linkage.bedpe')
#                                     
```
## Read in results from wynton
```{python}
adata_atac = sc.read_h5ad('../results_all/adata_atac.h5ad')
```

```{python}
plt.hist(adata_atac.X.sum(1), bins=100, range=(0, 100000))
plt.show()
```


```{python}
#barcodes from seurat object
adata_atac = adata_atac[adata_atac.obs_names.isin(barcodes)].copy()
adata_atac.obs_names
```
# We normalize aggregated peaks with TF-IDF.
```{python}
mv.tfidf_norm(adata_atac)
```

#Finding shared barcodes and features between RNA and ATAC
```{python}
shared_genes = pd.Index(np.intersect1d(adata_rna.var_names, adata_atac.var_names))
shared_cells = pd.Index(np.intersect1d(adata_rna.obs_names, adata_atac.obs_names))
len(shared_genes), len(shared_cells)
```
```{python}
adata_rna = adata_rna[shared_cells, shared_genes]
adata_atac = adata_atac[shared_cells, shared_genes]
```

```{python}
scv.pp.normalize_per_cell(adata_rna)
scv.pp.log1p(adata_rna)
scv.pp.moments(adata_rna, n_pcs=30, n_neighbors=50)
```
```{python}
sc.pl.umap(adata_rna, color=['annotation_broad'], frameon=False, legend_loc='on data')
```


```{python}
scv.pl.proportions(adata_rna, groupby='annotation_broad')
```
```{python}
scv.tl.velocity(adata_rna, mode='stochastic')
scv.tl.velocity_graph(adata_rna)
```
```{python}
scv.pl.velocity_embedding_grid(adata_rna, basis='umap', color='annotation_broad', save='embedding_grid.pdf', title='', scale=0.25)
scv.pl.velocity_embedding_stream(adata_rna, basis='umap', color='annotation_broad', save='embedding_stream.pdf', title='')
```
# Smoothing gene aggregagted peaks by neighbors
```{python}
nn_idx = np.loadtxt("../results_all/nn_idx.txt", delimiter=',')
nn_dist = np.loadtxt("../results_all/nn_dist.txt", delimiter=',')
nn_cells = pd.Index(pd.read_csv("../results_all/nn_cells.txt", header=None)[0])

np.all(nn_cells == adata_atac.obs_names)
```
```{python}
mv.knn_smooth_chrom(adata_atac, nn_idx, nn_dist)
```

# Running multi-omic dynamical model
```{python}
adata_result = mv.recover_dynamics_chrom(adata_rna,
                                         adata_atac,
                                         max_iter=5,
                                         init_mode="invert",
                                         parallel=True,
                                         save_plot=False,
                                         rna_only=False,
                                         fit=True,
                                         n_anchors=500,
                                         extra_color_key='annotation_broad'
                                        )
```


```{r}
library(reticulate)
table(py$suff)

```


