---
title: "8_rna_velocity"
---
```{r}
suppressPackageStartupMessages({
  library(Seurat)
  library(Signac)
  library(SeuratWrappers)
  library(BSgenome.Nfurzeri.NCBI.Nfu20140520.custom)
  library(tidyverse)
})



#import helper functions
source("C:/Users/Christopher He/OneDrive/UCSF/Singh Lab/diapause-sc-multiome/code/helper_functions.R")

#set output and figure directories
outs_dir <- "../results_all/"
fig_dir <- "../results_all/"
```
#Load Seurat Object
```{r}
# so <- readRDS("../results_all/so_neurons.rds")
```


```{r}
# so$barcode <- colnames(so)
# so$umap.rna_1 <- so@reductions$umap.rna@cell.embeddings[,1]
# so$umap.rna_2 <- so@reductions$umap.rna@cell.embeddings[,2]
# so$umap.atac_1 <- so@reductions$umap.atac@cell.embeddings[,1]
# so$umap.atac_2 <- so@reductions$umap.atac@cell.embeddings[,2]
# so$umap.wnn_1 <- so@reductions$wnn.umap@cell.embeddings[,1]
# so$umap.wnn_2 <- so@reductions$wnn.umap@cell.embeddings[,2]
# 
# # write metadata
# write.csv(so@meta.data, file=paste0(outs_dir, 'metadata.csv'), quote=F, row.names=F)
# 
# # write dimesnionality reduction matrix, in this example case pca matrix
# write.csv(so@reductions$pca@cell.embeddings, file=paste0(outs_dir, 'pca.csv'), quote=F, row.names=F)
# 
# # extract neighborhood graph
# nn_idx <- so@neighbors$weighted.nn@nn.idx
# nn_dist <- so@neighbors$weighted.nn@nn.dist
# nn_cells <- so@neighbors$weighted.nn@cell.names
# 
# # save neighborhood graph
# write.table(nn_idx, paste0(outs_dir, "nn_idx.txt"), sep = ',', row.names = F, col.names = F, quote = F)
# write.table(nn_dist, paste0(outs_dir, "nn_dist.txt"), sep = ',', row.names = F, col.names = F, quote = F)
# write.table(nn_cells, paste0(outs_dir, "nn_cells.txt"), sep = ',', row.names = F, col.names = F, quote = F)
# 
# rm(so)
# gc()
```


Note: Using python version through reticulate. scVelo and multiVelo were installed in the multiome-velocyto conda environment
```{r}
reticulate::use_condaenv("multiome-velocyto")
library(reticulate)
```

```{python}
import os
import scipy
import numpy as np
import pandas as pd
import scanpy as sc
import scvelo as scv
import multivelo as mv
import matplotlib.pyplot as plt
from pathlib import Path
```

```{python}
sc.settings.cachedir = Path('../results_all/cache')
scv.settings.verbosity = 3
scv.settings.presenter_view = True
scv.set_figure_params('scvelo')
pd.set_option('display.max_columns', 100)
pd.set_option('display.max_rows', 200)
np.set_printoptions(suppress=True)
```

## Reading in unspliced and spliced counts
```{python}
adata_rna = sc.read("../aggr_batch1-3/velocyto/velocyto_combined.loom", cache=True)

#rename barcodes to seurat format ("[barcode]-n" where n correspondes to sample)
suffixes = sum([[f"-{i}"] * n for i, n in enumerate([10177, 9774, 10510, 6575, 10736, 5993, 17082, 12245, 7928, 12501, 20000, 12563], start=1)], [])
adata_rna.obs_names = [x.split(':')[1][:-1] for x in adata_rna.obs_names]

#append "-n" suffix to each barcode
adata_rna.obs_names = [a + b for a, b in zip(adata_rna.obs_names, suffixes)]
```

```{python}
#read in metadata and pca table from seurat object
cell_meta = pd.read_csv("../results_all/metadata.csv")

#filter barcodes in seurat object
barcodes = cell_meta.barcode
adata_rna = adata_rna[adata_rna.obs_names.isin(barcodes)].copy()

#reorder metadata in order of adata
cell_meta = cell_meta.iloc[pd.Index(cell_meta['barcode']).get_indexer(adata_rna.obs_names)] 

#read in pca
pca = pd.read_csv("../results_all/pca.csv")
pca = pca.loc[cell_meta.index] # reorder pca in order of adata

#add metadata to adata
cell_meta = cell_meta.set_index('barcode')
adata_rna.obs = cell_meta


# set pca and umap
adata_rna.obsm['X_pca'] = pca.to_numpy()
adata_rna.obsm['X_umap'] = np.vstack((adata_rna.obs['umap.rna_1'].to_numpy(), adata_rna.obs['umap.rna_2'].to_numpy())).T
```


```{python}
#plot umap
sc.pl.umap(adata_rna, color=['timepoint_group'], frameon=False)
```

```{python}
# Top 2000 variable genes are used for downstream analyses.
scv.pp.filter_and_normalize(adata_rna, min_shared_counts=10, n_top_genes=2000)
```


##Preprocessing the ATAC counts (Done on wynton)
```{python}
# adata_atac = sc.read_10x_mtx('../aggr_batch1-3/filtered_feature_bc_matrix/', var_names='gene_symbols', cache=True, gex_only=False)
# adata_atac = adata_atac[:,adata_atac.var['feature_types'] == "Peaks"]
```

```{python}
# adata_atac = mv.aggregate_peaks_10x(adata_atac,
#                                     '../aggr_batch1-3/atac_peak_annotation.tsv',
#                                     '../aggr_batch1-3/feature_linkage.bedpe')
#                                     
```
## Read in results from wynton
```{python}
adata_atac = sc.read_h5ad('../results_all/adata_atac.h5ad')
```

```{python}
plt.hist(adata_atac.X.sum(1), bins=100, range=(0, 100000))
plt.show()
```


```{python}
#barcodes from seurat object
adata_atac = adata_atac[adata_atac.obs_names.isin(barcodes)].copy()
adata_atac.obs_names
```
# We normalize aggregated peaks with TF-IDF.
```{python}
mv.tfidf_norm(adata_atac)
```

#Finding shared barcodes and features between RNA and ATAC
```{python}
shared_genes = pd.Index(np.intersect1d(adata_rna.var_names, adata_atac.var_names))
shared_cells = pd.Index(np.intersect1d(adata_rna.obs_names, adata_atac.obs_names))
len(shared_genes), len(shared_cells)
```
```{python}
adata_rna = adata_rna[shared_cells, shared_genes]
adata_atac = adata_atac[shared_cells, shared_genes]
```

```{python}
#pre-process
scv.pp.normalize_per_cell(adata_rna)
scv.pp.log1p(adata_rna)
scv.pp.moments(adata_rna, n_pcs=30, n_neighbors=50)
```
```{python}
sc.pl.umap(adata_rna, color=['annotation_broad'], frameon=False, legend_loc='on data', palette = ['',])
sc.pl.umap(adata_rna, color=['timepoint_group'], frameon=False, legend_loc='on data', palette = ['#FD8D3C','#C6DBEF', '#2171B5', '#6BAED6', '#800080'])

```

# ScVelo

## Splicing proportions
```{python}
scv.pl.proportions(adata_rna, groupby='annotation_broad')
scv.pl.proportions(adata_rna, groupby='timepoint_group')

```
## Compute Velocities
```{python}
#compute velocity for each gene
scv.tl.velocity(adata_rna, mode='stochastic')

#compute transition probabilities using gene velocities
scv.tl.velocity_graph(adata_rna)
```

## Visualize velocity fields
```{python}
#visualize velocity fields
scv.pl.velocity_embedding_grid(adata_rna, basis='umap', color='annotation_broad', save='embedding_grid.pdf', title='', scale=0.25)
scv.pl.velocity_embedding_stream(adata_rna, basis='umap', color='annotation_broad', save='embedding_stream.pdf', title='')
```

## Plot some genes
```{python}
# plot velocity of a selected gene
scv.pl.velocity(adata_rna, var_names=['jade2'], color='timepoint_group')
scv.pl.velocity(adata_rna, var_names=['elavl3'], color='annotation_broad')

```

## Identify Important genes
```{python}
scv.tl.rank_velocity_genes(adata_rna, groupby='timepoint_group', min_corr=.3)

df = pd.DataFrame(adata_rna.uns['rank_velocity_genes']['names'])
df.head()
```
```{python}
kwargs = dict(frameon=False, size=10, linewidth=1.5, color = 'timepoint_group')

scv.pl.scatter(adata_rna, df['latedia'][:5], ylabel='latedia', **kwargs)
scv.pl.scatter(adata_rna, df['dev'][:5], ylabel='dev', **kwargs)
```

## Speed and coherence
Two more useful stats: - The speed or rate of differentiation is given by the length of the velocity vector. - The coherence of the vector field (i.e., how a velocity vector correlates with its neighboring velocities) provides a measure of confidence.
```{python}
scv.tl.velocity_confidence(adata_rna)
keys = 'velocity_length', 'velocity_confidence'
scv.pl.scatter(adata_rna, c=keys, cmap='coolwarm', perc=[5, 95])
```

## Velocity graph and pseudotime
```{python}
scv.pl.velocity_graph(adata_rna, threshold=.5, color = 'timepoint_group')
```

```{python}
scv.tl.velocity_pseudotime(adata_rna)
scv.pl.scatter(adata_rna, color='velocity_pseudotime', cmap='gnuplot')
```
## PAGA velocity graph
```{python}
!pip install python-igraph --upgrade --quiet
```


```{python}
# this is needed due to a current bug - bugfix is coming soon.
adata_rna.uns['neighbors']['distances'] = adata_rna.obsp['distances']
adata_rna.uns['neighbors']['connectivities'] = adata_rna.obsp['connectivities']

scv.tl.paga(adata_rna, groups='timepoint_group')
df = scv.get_df(adata_rna, 'paga/transitions_confidence', precision=2).T
df.style.background_gradient(cmap='Blues').format('{:.2g}')

df
```

```{python}
scv.pl.paga(adata_rna, basis='umap', size=50, alpha=.1, min_edge_width=2, node_size_scale=1.5)
```


# Multi-Velo
## Smoothing gene aggregagted peaks by neighbors
```{python}
nn_idx = np.loadtxt("../results_all/nn_idx.txt", delimiter=',')
nn_dist = np.loadtxt("../results_all/nn_dist.txt", delimiter=',')
nn_cells = pd.Index(pd.read_csv("../results_all/nn_cells.txt", header=None)[0])

np.all(nn_cells == adata_atac.obs_names)
```

```{python}
mv.knn_smooth_chrom(adata_atac, nn_idx, nn_dist)
```

## Running multi-omic dynamical model (done on wynton)
```{python}
adata_result = mv.recover_dynamics_chrom(adata_rna,
                                         adata_atac,
                                         max_iter=5,
                                         init_mode="invert",
                                         parallel=False,
                                         save_plot=False,
                                         rna_only=False,
                                         fit=True,
                                         n_anchors=500,
                                         extra_color_key='annotation_broad'
                                        )
```


```{r}
library(reticulate)
table(py$suff)

```


