---
title: "figure2"
---

```{r}
source("C:/Users/Christopher He/OneDrive/UCSF/Singh Lab/diapause-sc-multiome/code/helper_functions.R")

suppressPackageStartupMessages({
  library(scCustomize)
  library(ComplexHeatmap)
  library(circlize)
})



outs_dir <- "../results_all/"
fig_dir <- "../manuscript/figure2/"
```

```{r}
so <- readRDS("../results_all/so_wnn.rds")
```

# Global and Cell Type Genes 
```{r}
# read in master sheets
master_genes <- read.csv(paste0(outs_dir, "master_genes_global_and_celltype.csv"), row.names = 1)

# Compute Rank (number of significant cell types)
master_genes$rank <- apply(master_genes[,  grep("p_val_adj", names(master_genes), value = TRUE)], 1, function(row) sum(row < 0.05))
# celltypes to include
celltypes <- unique(sapply(colnames(master_genes %>% select(contains("pct.1"))), function(x) sub("pct.1_*", "", x)))
celltypes <- celltypes[-1] #remove global column
celltype_cols <- colors_broad[gsub("\\.", "/",celltypes)]
#column_cols[1] <- "#2171B5"
names(celltype_cols) <- celltypes

# Select global diapause genes
global_genes <- master_genes %>% filter(
  pct.1_global > 0.01 &
  p_val_adj_global < 0.05 &
  avg_log2FC_global > 1 &
  rank > 9
) %>% arrange(desc(avg_log2FC_global))

# global matrix
global <- as.matrix(global_genes %>% select(all_of(paste0("avg_log2FC_", celltypes))))
pval <- as.matrix(global_genes %>% select(all_of(paste0("p_val_adj_", celltypes))))
global_sig <- global * ifelse(pval < 0.05, 1, NA)
colnames(global_sig) <- celltypes
rownames(global_sig) <- global_genes$final_symbol


# select cell type diapause genes
dia_pct_thresh <- 0.1
p_val_adj_thresh <- 0.05
fc_thresh <- 1
rank_thresh <- 3

#rank each gene by number of cell types in which it is significantly DE
master_genes$rank <- rowSums(ifelse(master_genes %>% select(contains("p_val_adj")) < p_val_adj_thresh, 1, 0) == 1)

ctd_genes <- data.frame(avg_log2FC = numeric(), gene = character(), celltype = character())
for(celltype in celltypes[2:length(celltypes)]){
  genes <- master_genes %>% filter(
    .data[[paste0("pct.1_", celltype)]] > dia_pct_thresh &
    .data[[paste0("p_val_adj_", celltype)]] < p_val_adj_thresh &
    .data[[paste0("avg_log2FC_", celltype)]] > fc_thresh &
    rank < rank_thresh
  ) %>% arrange(desc(.data[[paste0("avg_log2FC_", celltype)]])) %>% select(paste0("avg_log2FC_", celltype))
  
  if(nrow(genes) > 0){
    genes$gene <- rownames(genes)
    rownames(genes) <- NULL
    genes$celltype <- celltype
    colnames(genes) <- colnames(ctd_genes)
    ctd_genes <- rbind(ctd_genes, genes)
  }
}

#genes to include
ctd_genes <- ctd_genes %>%
  group_by(gene) %>%
  filter(avg_log2FC == max(avg_log2FC)) %>%
  ungroup()

#ctd matrix
ctd_FC <- as.matrix(master_genes[ctd_genes$gene,] %>% select(all_of(paste0("avg_log2FC_", celltypes))))
ctd_pval <- as.matrix(master_genes[ctd_genes$gene,] %>% select(all_of(paste0("p_val_adj_", celltypes))))

ctd_FC_sig <- ctd_FC * ifelse(ctd_pval < 0.05, 1, NA)
colnames(ctd_FC_sig) <- celltypes
rownames(ctd_FC_sig) <- translate(ctd_genes$gene)

## Plot merged genes
merged_FC_sig <- as.matrix(bind_rows(as.data.frame(global_sig), as.data.frame(ctd_FC_sig)))


ct_colors = rowAnnotation(ct = colnames(merged_FC_sig), col = list(ct = celltype_cols), show_legend=F, show_annotation_name=F)
ctg_colors = HeatmapAnnotation(
  ctg = c(rep("global", nrow(global_sig)), ctd_genes$celltype),
  col = list(ctg = c(column_cols, "global" = "#2171B5")),
  show_legend = F, show_annotation_name = F, simple_anno_size = unit(0.3, "cm"))


hm <- Heatmap(
  t(merged_FC_sig),
  col = colorRamp2(breaks = c(-5,0,0.1,5,7), colors = c("blue","white", "pink", "red", "red")),
  na_col = "grey90",
  cluster_rows = F, 
  cluster_columns = F,
  show_column_names = F,
  rect_gp = gpar(col = "white", lwd = 0.5),
  right_annotation = ct_colors,
  top_annotation = ctg_colors,
  heatmap_legend_param = list(title = "Fold Change in Diapause, adjusted p-val < 0.05", legend_direction = "horizontal"),
  width = unit(0.2*nrow(merged_FC_sig), "cm"),
  height = unit(0.5*ncol(merged_FC_sig), "cm"),
  column_split = c(rep("(1) global",nrow(global_sig)), rep("(2) ctd",nrow(ctd_FC_sig)))
)

SaveFigure(hm, type = "png", "merged_global_ctd_gene_heatmap", height = 5, width = 20)

```

# Global and Cell Type Peaks 
```{r}
# read in master sheets
master_peaks <- read.csv(paste0(outs_dir, "master_peaks_global_and_celltype.csv"), row.names = 1)
rownames(master_peaks) <- master_peaks$Row.names

# Compute Rank (number of significant celltypes)
master_peaks$rank <- apply(master_peaks[,  grep("p_val_adj", names(master_peaks), value = TRUE)], 1, function(row) sum(row < 0.05))

# celltypes to include
celltypes <- unique(sapply(colnames(master_genes %>% select(contains("pct.1"))), function(x) sub("pct.1_*", "", x)))
celltypes <- celltypes[-1] #remove global column
celltype_cols <- colors_broad[gsub("\\.", "/",celltypes)]
#column_cols[1] <- "#2171B5"
names(celltype_cols) <- celltypes

#global peaks
global_peaks <- master_peaks %>% filter(
  pct.1_global > 0.01 &
  p_val_adj_global < 0.05 &
  avg_log2FC_global > 1 &
  rank > 9
) %>% arrange(desc(avg_log2FC_global))


#global matrix
global <- as.matrix(global_peaks %>% select(all_of(paste0("avg_log2FC_", celltypes))))
pval <- as.matrix(global_peaks %>% select(all_of(paste0("p_val_adj_", celltypes))))
global_sig <- global * ifelse(pval < 0.05, 1, NA)
colnames(global_sig) <- celltypes


dia_pct_thresh <- 0.1
p_val_adj_thresh <- 0.05
fc_thresh <- 1
rank_thresh <- 3

#rank each gene by number of cell types in which it is significantly DE
master_peaks$rank <- rowSums(ifelse(master_peaks %>% select(contains("p_val_adj")) < p_val_adj_thresh, 1, 0) == 1)

ctd_peaks <- data.frame(avg_log2FC = numeric(), peak = character(), celltype = character())
for(celltype in celltypes[2:length(celltypes)]){
  peaks <- master_peaks %>% filter(
    .data[[paste0("pct.1_", celltype)]] > dia_pct_thresh &
    .data[[paste0("p_val_adj_", celltype)]] < p_val_adj_thresh &
    .data[[paste0("avg_log2FC_", celltype)]] > fc_thresh &
    rank < rank_thresh
  ) %>% arrange(desc(.data[[paste0("avg_log2FC_", celltype)]])) %>% select(paste0("avg_log2FC_", celltype))
  
  if(nrow(peaks) > 0){
    peaks$peak <- rownames(peaks)
    rownames(peaks) <- NULL
    peaks$celltype <- celltype
    colnames(peaks) <- colnames(ctd_peaks)
    ctd_peaks <- rbind(ctd_peaks, peaks)
  }
}

#peaks to include
ctd_peaks <- (ctd_peaks %>%
  group_by(peak) %>%
  filter(avg_log2FC == max(avg_log2FC)) %>%
  ungroup())


ctd_FC <- as.matrix(master_peaks[ctd_peaks$peak,] %>% select(all_of(paste0("avg_log2FC_", celltypes))))
ctd_pval <- as.matrix(master_peaks[ctd_peaks$peak,] %>% select(all_of(paste0("p_val_adj_", celltypes))))

ctd_FC_sig <- ctd_FC * ifelse(ctd_pval < 0.05, 1, NA)
colnames(ctd_FC_sig) <- celltypes

#plot merged heatmap
merged_FC_sig <- as.matrix(bind_rows(as.data.frame(global_sig), as.data.frame(ctd_FC_sig)))


ct_colors = rowAnnotation(ct = colnames(merged_FC_sig), col = list(ct = celltype_cols), show_legend=F, show_annotation_name=F)
ctg_colors = HeatmapAnnotation(
  ctg = c(rep("global", nrow(global_sig)), ctd_peaks$celltype),
  col = list(ctg = c(celltype_cols, "global" = "#2171B5")),
  show_legend = F, show_annotation_name = F, simple_anno_size = unit(0.3, "cm"))


hm <- Heatmap(
  t(merged_FC_sig),
  col = colorRamp2(breaks = c(-5,0,0.1,5,7), colors = c("blue","white", "pink", "red", "red")),
  na_col = "grey90",
  cluster_rows = F, 
  cluster_columns = F,
  show_column_names = F,
  rect_gp = gpar(col = "white", lwd = 0.5),
  right_annotation = ct_colors,
  top_annotation = ctg_colors,
  heatmap_legend_param = list(title = "Fold Change in Diapause, adjusted p-val < 0.05", legend_direction = "horizontal"),
  width = unit(0.2*nrow(merged_FC_sig), "cm"),
  height = unit(0.5*ncol(merged_FC_sig), "cm"),
  column_split = c(rep("(1) global",nrow(global_sig)), rep("(2) ctd",nrow(ctd_FC_sig)))
)

SaveFigure(hm, type = "png", "merged_global_ctd_peak_heatmap", height = 5, width = 40)

```
# Global and Cell Type Motifs
```{r}
master_motifs_ct <- read.csv(paste0(outs_dir, "master_motifs_global_and_celltype.csv"))

ct_motifs <- master_motifs_ct[, c(2,16:44)]
ct_motifs$group <- factor(ct_motifs$group, levels = c("global", celltypes))
ct_motifs <- ct_motifs %>% filter(!is.na(group)) %>% arrange(group)


ct_motif_mat <- as.matrix(ct_motifs %>% column_to_rownames(var = "motif.name.x") %>% select(contains("fold.enrichment")))
colnames(ct_motif_mat) <- gsub("fold.enrichment_", "", colnames(ct_motif_mat))

columns <- colnames(ct_motif_mat)
column_cols <- colors_broad[gsub("\\.", "/",columns)]
names(column_cols) <- columns


# motif_names <- rownames(ct_motif_mat)
# motif_names[-seq(1, length(motif_names), by = 5)] <- ""
# rownames(ct_motif_mat) <- motif_names

ct_colors = rowAnnotation(ct = columns, col = list(ct = column_cols), show_legend=F, show_annotation_name=F)
col_colors = HeatmapAnnotation(celltype = ct_motifs$group, col = list(celltype = c(global = "#2171B5", column_cols)), show_legend = F, show_annotation_name = F, simple_anno_size = unit(0.3, "cm"))


hm <- Heatmap(
  t(ct_motif_mat),
  col = colorRamp2(breaks = c(0,3), colors = c("white",  "red")),
  na_col = "grey90",
  cluster_rows = F, 
  cluster_columns = F,
  show_column_names = F,
  rect_gp = gpar(col = "white", lwd = 0.5),
  right_annotation = ct_colors,
  top_annotation = col_colors,
  heatmap_legend_param = list(title = "Fold Enrichment\n in Diapause,\n adjusted p-val < 0.05", legend_direction = "horizontal"),
  column_split = c(rep("(1) global",sum(ct_motifs$group == "global")), rep("(2) ctd", sum(ct_motifs$group != "global")))
)

hm <- draw(hm, heatmap_legend_side = "bottom")
SaveFigure(hm, type = "png", "ct_motif_heatmap", height = 5, width = 25)

```



#Example Genes
## jade2
```{r}
DefaultAssay(so) <- "RNA"
FeaturePlot(so, "jade2", reduction = "umap") + NoAxes()

Idents(so) <- "annotation_broad"
p <- VlnPlot(so, "jade2", group.by = "annotation_broad", split.by = "devtraj", idents = gsub("\\.", "/",celltypes), cols =c("#800080", "#FD8D3C", "#6BAED6"), pt.size = 0)

SaveFigure(p, "jade2_opt1", type = "png", width = 8, height = 5, res = 300)


p1 <- p + geom_jitter(mapping = aes(color = split), data = p$data,
  position = position_jitterdodge(jitter.width = 0.4, dodge.width = 0.9), size = 0.5, alpha = 0.5)+
  scale_color_manual(breaks = c("pre-dia","dev", "dia"), values = c("#800080", "#FD8D3C", "#6BAED6"))


SaveFigure(p1, "jade2_opt2", type = "png", width = 8, height = 5, res = 300)

p2 <- p + geom_jitter(mapping = aes(color = split), data = p$data,
  position = position_jitterdodge(jitter.width = 0.4, dodge.width = 0.9), size = 0.5)+
  scale_color_manual(breaks = c("pre-dia","dev", "dia"), values = c("#e895e8", "#ffb885", "#a4d2ed"))


SaveFigure(p2, "jade2_opt3", type = "png", width = 8, height = 5, res = 300)


##TODO: add star for significance
##TODO: change to same groups as heatmap
```
## sptssb
```{r}
DefaultAssay(so) <- "RNA"
FeaturePlot(so, "sptssb", reduction = "umap") + NoAxes()

Idents(so) <- "annotation_broad"
p <- VlnPlot(so, "sptssb", group.by = "annotation_broad", split.by = "devtraj", idents = gsub("\\.", "/",celltypes), cols =c("#800080", "#FD8D3C", "#6BAED6"), pt.size = 0)

SaveFigure(p, "sptssb_opt1", type = "png", width = 8, height = 5, res = 300)


p1 <- p + geom_jitter(mapping = aes(color = split), data = p$data,
  position = position_jitterdodge(jitter.width = 0.4, dodge.width = 0.9), size = 0.5, alpha = 0.5)+
  scale_color_manual(breaks = c("pre-dia","dev", "dia"), values = c("#800080", "#FD8D3C", "#6BAED6"))


SaveFigure(p1, "sptssb_opt2", type = "png", width = 8, height = 5, res = 300)

p2 <- p + geom_jitter(mapping = aes(color = split), data = p$data,
  position = position_jitterdodge(jitter.width = 0.4, dodge.width = 0.9), size = 0.5)+
  scale_color_manual(breaks = c("pre-dia","dev", "dia"), values = c("#e895e8", "#ffb885", "#a4d2ed"))


SaveFigure(p2, "sptssb_opt3", type = "png", width = 8, height = 5, res = 300)
p2
```
##All genes
```{r}
so$rna.splits <- ifelse(so$timepoint %in% c("hb", "dev1d"), "dev", 
                                               ifelse(so$timepoint %in% c("dia6d", "dia1mo"), "dia", NA))
so_rna <- subset(so, subset = rna.splits %in% c("dev", "dia"))
```


```{r}
DefaultAssay(so) <- "RNA"

for(i in 1:nrow(ctd_genes)){
  row = ctd_genes[i,]
  Idents(so_rna) <- "annotation_broad"
  p <- VlnPlot(
    so_rna,
    row$gene,
    group.by = "annotation_broad",
    split.by = "rna.splits",
    idents = gsub("\\.", "/",celltypes),
    cols =c("#FD8D3C", "#6BAED6"),
    pt.size = 0
  )
  p1 <- p + geom_jitter(mapping = aes(color = split), data = p$data,
                        position = position_jitterdodge(jitter.width = 0.4, dodge.width = 0.9), size = 0.5, alpha = 0.2) +
    scale_color_manual(breaks = c("dev", "dia"), values = c("#FD8D3C", "#6BAED6"))

  SaveFigure(p1, paste0("ctd_genes/", i, "_", row$gene, "_", row$celltype), type = "png", width = 8, height = 5, res = 300)
}
```


# Example Peaks
## function
```{r}
plot_dia_vs_dev_peak <- function(
    so,
    peak,
    ident,
    tp1,
    tp2,
    tp1.name = "dia",
    tp1.color = "#6BAED6",
    tp2.name = "dev",
    tp2.color = "#FD8D3C",
    extend.upstream = 5000,
    extend.downstream = 5000,
    ymax = 100
){
  DefaultAssay(so) <- "ATAC"
  
  #create metadata column for timepoints
  so$tp <- ifelse(so$timepoint %in% tp1, tp1.name, 
                  ifelse(so$timepoint %in% tp2, tp2.name, "other"))
  so$tp <- factor(so$tp, levels = c(tp1.name, tp2.name, "other"))
  
  #subset object by cell type given
  Idents(so) <- "annotation_broad"
  so_sub <- subset(so, idents = ident)
  l <- levels(so_sub$tp)
  so_sub$tp <- paste0(so_sub$tp, "_", ident)
  so_sub$tp <- factor(so_sub$tp, levels = paste0(l, "_", ident))

  #create coverage plot for cell type subset
  Idents(so_sub) <- "tp"
  ct_cov <- CoveragePlot(
    object = so_sub,
    region = peak,
    annotation = F,
    peaks = F,
    links = F,
    idents = levels(so_sub$tp)[1:2],
    extend.upstream = extend.upstream,
    extend.downstream = extend.downstream,
    ymax = ymax
  ) & scale_fill_manual(values = c(tp1.color, tp2.color))
  return(ct_cov)
}

```

##All Peaks
```{r}
DefaultAssay(so) <- "ATAC"

for(rownum in 19:nrow(ctd_peaks)){
  row = ctd_peaks[rownum,]
  cts <- gsub("\\.", "/",celltypes)
  plots <- list()
  for(i in 1:length(cts)){
    plot <- plot_dia_vs_dev_peak(
      so,
      peak = row$peak,
      ident = gsub("\\.", "/", cts[i]), 
      tp1 = c("dia6d", "dia1mo"),
      tp2 = c("hb", "dev1d"),
      ymax = 100
    )
    plots[[i]] <- plot
  }
  gene_plot <- AnnotationPlot(
    so,
    region = row$peak,
    extend.upstream = 5000,
    extend.downstream = 5000
  )
  plots[[i+1]] <- gene_plot
  p <- CombineTracks(
    plotlist = plots,
    heights = rep(1, 9)
  )
  
  
  SaveFigure(p, type = "png", paste0("ctd_peaks/", rownum, "_", row$peak, "_", row$celltype), height = 8, width = 5)
}

``` 


#Pseudobulking
```{r}
pseudo <- AggregateExpression(so, assays = "RNA", return.seurat = T, group.by = c("orig.ident", "annotation_broad"))

pseudo_cells <- Cells(pseudo)

dev_somite_cells <- pseudo_cells[grep("^(hb|dev1d).*_somite$", pseudo_cells)]
dia_somite_cells <- pseudo_cells[grep("^(dia6d|dia1mo).*_somite$", pseudo_cells)]

pseudobulk.de <- FindMarkers(
  object = pseudo,
  ident.1 = dia_somite_cells,
  ident.2 = dev_somite_cells,
)


```

```{r}
pseudobulk.de
```

#Peak Analysis
Category	Definition
Opened	Low early, significantly increased accessibility later
Closed	High early, significantly decreased later
Maintained	No significant change across timepoints
Dynamically Regulated	Any significant non-monotonic or timepoint-specific accessibility pattern

Method: peak by timepoint matrix

```{r}
DefaultAssay(so) <- "ATAC"

# Get the raw peak x cell matrix
peak_matrix <- GetAssayData(so, layer = "counts")

# Timepoint metadata (replace "timepoint" with your actual column name)
timepoints <- so$timepoint

# How many cells per timepoint? Choose the **minimum** across all timepoints
min_cells <- min(table(timepoints))

# Unique timepoints
unique_tp <- unique(timepoints)

# Initialize pseudobulk matrix
pseudobulk_matrix <- matrix(0, nrow = nrow(peak_matrix), ncol = length(unique_tp))
rownames(pseudobulk_matrix) <- rownames(peak_matrix)
colnames(pseudobulk_matrix) <- unique_tp

# For each timepoint: randomly subsample 'min_cells' and sum counts
set.seed(123)  # for reproducibility

for (tp in unique_tp) {
  cells_tp <- colnames(so)[timepoints == tp]
  
  # Randomly sample min_cells
  sampled_cells <- sample(cells_tp, size = min_cells, replace = FALSE)
  
  # Sum the counts across sampled cells
  pseudobulk_matrix[, tp] <- Matrix::rowSums(peak_matrix[, sampled_cells, drop = FALSE])
}


# scale
zscore_matrix <- t(scale(t(pseudobulk_matrix)))
zscore_matrix_dia <- zscore_matrix[,c("hb", "dia1d", "dia2d", "dia6d", "dia1mo", "dia3mo")]

# top variable peaks
variable_peaks <- apply(zscore_matrix_dia, 1, sd)
top_peaks <- order(variable_peaks, decreasing = TRUE)[1:5000]
top_peaks <- zscore_matrix_dia[top_peaks, ]
```
## Clustering options
```{r}
library(RColorBrewer)
k <- 6
cluster_colors <- setNames(brewer.pal(k, "Set2"), as.character(1:k))
```

### hierarchical
```{r}
# Distance: Euclidean; Method: complete linkage
dist_matrix <- dist(top_peaks, method = "euclidean")
hclust_res <- hclust(dist_matrix, method = "complete")

# Cut into k clusters (e.g., 6 clusters)
k <- 6
peak_clusters_hclust <- cutree(hclust_res, k = k)


# Visualize with a heatmap
library(pheatmap)
pheatmap(top_peaks[order(peak_clusters_hclust),], cluster_rows = FALSE, cluster_cols = FALSE, show_rownames = FALSE,
         annotation_row = data.frame(Cluster = factor(peak_clusters_hclust)), annotation_colors = list(Cluster = cluster_colors))
```

### k-means
```{r}
set.seed(123)  # for reproducibility
k <- 6
kmeans_res <- kmeans(top_peaks, centers = k, nstart = 10)

peak_clusters_kmeans <- kmeans_res$cluster

# Heatmap with cluster annotations
pheatmap(top_peaks[order(peak_clusters_kmeans),], cluster_rows = FALSE, cluster_cols = FALSE, show_rownames = FALSE,
         annotation_row = data.frame(Cluster = factor(peak_clusters_kmeans)), annotation_colors = list(Cluster = cluster_colors))

```

### Fuzzy C-means
```{r}
# Install and load the e1071 or Mfuzz package
install.packages("e1071")  # OR install.packages("Mfuzz")
library(e1071)

# Use cmeans from e1071
k <- 6
fcm_res <- cmeans(zscore_matrix, centers = k, m = 2)

# Peak membership scores and max cluster assignment
peak_membership <- fcm_res$membership
peak_clusters_fuzzy <- apply(peak_membership, 1, which.max)

# Optional: filter on high-confidence memberships
confident_peaks <- apply(peak_membership, 1, max) > 0.7

# Heatmap
pheatmap(zscore_matrix[confident_peaks, ], cluster_rows = TRUE, cluster_cols = FALSE,
         show_rownames = FALSE,
         annotation_row = data.frame(Cluster = factor(peak_clusters_fuzzy[confident_peaks])))
```

```{r}
# Prepare long-form dataframe
z_df <- as.data.frame(top_peaks)
z_df$Cluster <- factor(peak_clusters_hclust)
z_df$peak <- rownames(z_df)

long_df <- pivot_longer(
  z_df, 
  cols = -c(Cluster, peak),
  names_to = "Timepoint", 
  values_to = "Zscore"
)

# Ensure Timepoint is ordered correctly (if it's character)
long_df$Timepoint <- factor(long_df$Timepoint, levels = c("hb", "dia1d", "dia2d", "dia6d", "dia1mo", "dia3mo"))

# Summarize mean Z-score per cluster and timepoint
summary_df <- long_df %>%
  group_by(Cluster, Timepoint) %>%
  summarise(meanZ = mean(Zscore, na.rm = TRUE), .groups = "drop")

# Plot with facets
ggplot(summary_df, aes(x = Timepoint, y = meanZ, group = Cluster)) +
  geom_line(aes(color = Cluster), size = 1.2) +
  facet_wrap(~ Cluster, scales = "fixed") +  # One panel per cluster
  scale_color_manual(values = cluster_colors) +
  theme_minimal(base_size = 14) +
  labs(
    title = "Temporal Accessibility Dynamics per Cluster",
    y = "Mean Z-score",
    x = "Timepoint"
  ) +
  theme(legend.position = "none") + theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
```

